---
title: "Oviposition Logistic"
author: "Ryan Peek"
date: "Updated: `r format(Sys.Date())`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = normalizePath("../"))
```

# What Permits Plasticity in Breeding Timing in River Frogs?

How might varying environmental conditions provide cues for breeding timing from year to year? For frogs that have an amazing plasticity in breeding oviposition timing (inital deposition of eggs) such as *Rana boylii*, there must be certain thresholds that exist (a cutoff which makes oviposition not possible, physiologically). However, beyond a given threshold, there also may be certain cues which act as important environmental markers in dynamic environments such as rivers. By utilizing both, it is likely that the odds of reproductive success improve, and over time this results in evolutionary success. For example, frogs that breed to soon are more likely to have eggs scoured out by late spring storms or have longer developmental/hatching times because water temperatures are lower. Frogs breeding later in the season may trade off greater egg deposition and hatching success with later metamorphosis and lower overwinter survival. 

We know that frogs from the same populations in the exact same locations can deposit eggs more than a month later or many weeks earlier than the average observed date of deposition, depending on the annual conditions (i.e., wetter or drier years). We suspect that the water temperature may be the most important driver in this plasticity, as it may act as a threshold which once exceeded, provides a very strong indicator of "summer" or future hydrologic condition.

However, the flow in a given reach of river can also be highly correlated with water temperature. Linkage between water temperature and patterns in flow may be tightly tied, particularly in rivers in the Sierra Nevada. Historically, Sierra Nevada river hydrology has been driven by a general Mediterranean pattern of wet (though variable) winter weather followed by spring snowmelt and warmer and drier summers. While temperature may exist as a spawning threshold, the actual initiation of breeding or spawning is highly variable in **R. boylii**, and likely requires additional enviromental cues such as changing water levels, flow magnitude, air temperature, humidity, etc.

These additional metrics may each act as environmental cues, but to conserve plasticity across environmental variablity present in a wide geographic range, it is unlikely that any of these can dictate oviposition independently (whereas a water temperature threshold may be a sole driver in some years). In order to determine the strongest predictor of breeding, we've collected oviposition data over 6 rivers in the Sierra Nevada from 2011 through 2016. In addition we collected a variety of metrics related to flow, air temperature, water temperature, water year index, day of water year, precipitation, and barometric pressure.

## Try Model, Model Fail

Let's try to assess whether water temperature, air temperature, flow (daily/weekly recession rates & discharge) play a role in predicting when breeding (oviposition occurs). To build this model we need to block by river, and by water year, as these are independent (though water year isn't *fixed*, it is fixed across all sites per each year).

First let's take a look at some test data. Using a subset (say just pull the **NF American**), can we build a model that gets at whether there is predictability in these variables despite shifts in timing and magnitude from year to year? Ideally this seems to be setup for a logistic regression, where 1 is the breeding initation or observation of eggs in the river, and 0 is all the days prior to that event. An *Event History Analysis* or *Survival Analysis* approach might work (see section on geometric distribution, Ch10, pg 328). 

### Get the Data & Tidy

```{r get and tidy data, eval=F, echo=F}

library(tidyverse)
library(lubridate)

load("data/master_dat_2011-2016.rda") 
load("data/flow_dv_cfs_2011_6sites.rda") # updated and merged flows:

df <- master_df
# df <- master_df %>% filter(site!="MFY")
# df <- master_df %>% filter(month(date)>2 & month(date)<8 & !site=="MFA")
# df <- master_df %>% filter(month(date)>3 & month(date)<8 & site=="NFA" | site=="NFY")


# 2 NAs in dataset from 30 day avgs
df[!is.na(df$missData) & df$site=="RUB" & df$date==ymd("2011-05-26"),]$temp_30_max<-10.99
df[!is.na(df$missData) & df$site=="RUB" & df$date==ymd("2011-05-26"),]$temp_30_min<-9.7

rabo <- df[!is.na(df$missData),] %>% arrange(date)

# add unique rownames based on siteID
rabo <- rabo %>%
  mutate("siteID" = paste0(site, "-", row.names(.))) %>% 
  column_to_rownames(var = "siteID")  %>% as.data.frame
#names(rabo)
#rownames(rabo)

# create column with date for 15 and 30 day lags (for binomial logistic)
rabo_lag <- rabo %>% 
  mutate(d07 = date - 7, 
         d14 = date - 14,
         d21 = date - 21,
         d30 = date - 30) %>% 
  select(site, date, WY, d07:d30)

# rejoin / filter form orig dataset

rabo_bin7<- inner_join(df, rabo_lag, by=c("site"="site", "WY"="WY", "date"="d07")) %>% select(-c(date.y, d14:d30)) %>% mutate(lagEM="d07")
rabo_bin14 <- inner_join(df, rabo_lag, by=c("site"="site", "WY"="WY", "date"="d14")) %>% select(-c(date.y, d07:d30)) %>% mutate(lagEM="d14")
rabo_bin21 <- inner_join(df, rabo_lag, by=c("site"="site", "WY"="WY", "date"="d21")) %>% select(-c(date.y, d07:d30)) %>% mutate(lagEM="d21")
rabo_bin30 <- inner_join(df, rabo_lag, by=c("site"="site", "WY"="WY", "date"="d30")) %>% select(-c(date.y, d07:d21)) %>% mutate(lagEM="d30")

# bind up for one dataset of lags, add 1/0 col for breeding
rabo_LAGS<-rbind(rabo_bin7, rabo_bin14, rabo_bin21, rabo_bin30) %>% 
  mutate(breed = 0)
rm(list=ls(pattern = "rabo_bin*"))

# add breed col to orig dataset:
rabo <- rabo %>% mutate(breed = 1, lagEM="d00")

# make final logistic dataset, add 1/0 column
rabo_logist <- rbind(rabo_LAGS, rabo)
#save(rabo_logist, file = "models/rabo_logistic_all_sites.rda")
```

Prepare data by removing much of the information that will be unused, and converting and scaling the data as needed.

```{r prepData, eval=T, echo=T}

library(tidyverse)
library(lubridate)
library(rethinking)

load("models/rabo_logistic_all_sites.rda")

# Filter down data, use only unreg sites
d1 <- dplyr::select(rabo_logist, site, breed, lagEM, DOWY, everything(), 
                    -date, -EM_per_km, -REG, -(obs_strt:apr_jul),
                    -len_km, -starts_with("CDEC"), -station, -DOY,
                    -WYsum, -lev_7_avg, -Q_cfs, -Q_min, -Q_7_cfs) %>% 
    #filter(!site=="MFA") %>% 
  filter(site=="NFY" | site=="NFA") %>%
  as.data.frame

# RM NAs, add id and site to rownames:
d1 <- d1 %>% filter(!is.na(temp_30_min))
d1_rownames<- paste0(d1$site, "-",d1$WY, "-", seq(1:nrow(d1)))
d1$site <- as.factor(d1$site)
d1$lagEM <- as.factor(d1$lagEM)

# custom scale function
cusScale <- function(x){
  (x - mean(x))/sd(x)
}

unScale <- function(x, y){
  (x * sd(y) + mean(y))
}

# scale all data and rebind site and breed info:
d1s <- d1
d1s[,c(4:29)] <- apply(d1[,c(4:29)], 2, cusScale) # center and scale
d1s$site <- coerce_index(d1s$site) # for modeling
d1s$WY <- coerce_index(d1s$WY) # for modeling
d1s$lagEM <- coerce_index(d1s$lagEM) # for modeling
d1s <- d1s %>% rename(river=site, wyind=Index) # rename some vars

```

## Binomial (Logistic) Regression: Temp Only, No varying intercept

First let's create a model with all temperature data in it (*30-day, 7-day, and CV*). 

```{r temp_1.1a}

# just water temperature
m1.1a <- map2stan(
    alist(
        breed ~ dbinom( 1 , p ) ,
        logit(p) <- a + bt7mx*temp_7_max + bt7av*temp_7_avg + bt7mn*temp_7_min +
           btCV*temp_CV + bt30mx*temp_30_max + bt30mn*temp_30_min,
        a ~ dnorm(0,1),
        c(bt7mx, bt7av, bt7mn, btCV, bt30mx, bt30mn) ~ dnorm(0,1)
    ), data=d1s, warmup=1000, iter=5000, chains=4, cores=2)

precis(m1.1a)
plot(m1.1a)
dev.off()
logistic(0.65) # temp 7 max
logistic(0.56) # temp30 min (only 2% lower in prob for mean)

plot(precis(m1.1a))

# show density of probabilities
post1a <- extract.samples(m1.1a)
dens( post1a$bt7mx,  ylim=c(0,1.5), xlim=c(-5, 5))
dens( post1a$bt7av, col="gray20", lty=2, add=TRUE)
dens( post1a$bt7mn , col=rangi2, add=TRUE)
dens( post1a$bt30mx, col="red2" , lwd=2 , add=TRUE )
dens( post1a$bt30mn, col="maroon", lty=4 , add=TRUE )
text( 0.4 , 1 , "30dmn" , col="maroon" )
text( 2 , 0.4 , "7dmx" )

# diff between distributions of 30dmn and 7dmx
mu.1a.7mx <- logistic(post1a$a + post1a$bt7mx)
mu.1a.30mn <- logistic(post1a$a + post1a$bt30mn)
precis(data.frame(mu.1a.30mn,mu.1a.7mx))
dens(mu.1a.30mn, xlim=c(0, 1), show.HPDI = T)
dens(mu.1a.7mx, col="red2", add=T, show.HPDI=T)

# difference 
diff_30mn_7mx <- mu.1a.7mx - mu.1a.30mn
quantile( diff_30mn_7mx , probs=c(0.025,0.5,0.975) )

```

This model has the main variables that seem strongest, (*7-day max, 30 day min, and CV of temp*).

```{r temp_1.1b}
# now just 7day max and 30dymin
m1.1b <- map2stan(
    alist(
        breed ~ dbinom( 1 , p ) ,
        logit(p) <- a + bt7mx*temp_7_max + bt30mn*temp_30_min + btCV*temp_CV ,
        a ~ dnorm(0,1),
        c(bt7mx, bt30mn, btCV) ~ dnorm(0,1)
    ), data=d1s, warmup=1000, iter=5000, chains=4, cores=2)

plot(m1.1b)
dev.off()
precis(m1.1b)
plot(precis(m1.1b))

# show density of probabilities
post1b <- extract.samples(m1.1b)
dens( post1b$bt7mx , ylim=c(0,1.5), xlim=c(-1.4,3.4))
dens( post1b$bt30mn, col=rangi2 , lwd=2 , add=TRUE )
dens( post1b$btCV , col="red", add=TRUE)

# diff between distributions of 30dmn and 7dmx
mu.1b.7mx <- logistic(post1b$a + post1b$bt7mx)
mu.1b.30mn <- logistic(post1b$a + post1b$bt30mn)
mu.1b.CV <- logistic(post1b$a + post1b$btCV)
precis( data.frame(mu.1b.CV,mu.1b.7mx, mu.1b.30mn) )

# difference 
diff_CV_7mx <- mu.1b.7mx - mu.1b.CV
diff_30mn_7mx <- mu.1b.30mn - mu.1b.7mx
quantile( diff_30mn_7mx , probs=c(0.025,0.5,0.975) )
quantile( diff_CV_7mx , probs=c(0.025,0.5,0.975) )


```

And this includes a few other variables like *precipitation*.

```{r temp_1.1c}

# now add days no rain and top vars from last two
m1.1c <- map2stan(
    alist(
        breed ~ dbinom( 1 , p ) ,
        logit(p) <- a + bt7mx*temp_7_max + bt30mn*temp_30_min + btCV*temp_CV + bppt*days_no_ppt ,
        a ~ dnorm(0,1),
        c(bt7mx, bt30mn, btCV, bppt) ~ dnorm(0,1)
    ), data=d1s, warmup=1000, iter=5000, chains=4, cores=2)

plot(m1.1c)
precis(m1.1c)
dev.off()
plot(precis(m1.1c))

# show density of probabilities
post1c <- extract.samples(m1.1c)
dens( post1c$bt7mx , ylim=c(0,1.7), xlim=c(-1,2.4), show.HPDI = 0.9, show.zero = T)
text(1.5, 1, "7-day-max")
dens( post1c$bt30mn, col="maroon" , lty=2 , add=TRUE, show.HPDI = 0.9)
text(.5, 1.2, "30-day-min", col="maroon")
dens( post1c$btCV, col=rangi2 , lwd=2 , add=TRUE )
text(-0.4, 0.8, "temp_CV", col=rangi2)
dens( post1c$bppt , col="red", add=TRUE)
text(-0.65, 1.2, "days_no_ppt", col="red")

# compute diff between 30min and 7dymax
mu.1c.7mx <- post1c$a + post1c$bt7mx
mu.1c.30mn <- post1c$a + post1c$bt30mn
mu.1c.tCV <- post1c$a + post1c$btCV


# difference 
diff_7mx_30mn <- mu.1c.7mx - mu.1c.30mn
diff_7mx_CV <- mu.1c.7mx - mu.1c.tCV
quantile( diff_7mx_CV , probs=c(0.025,0.5,0.975) )
quantile( diff_7mx_30mn , probs=c(0.025,0.5,0.975) )
```

Let's compare all three *temperature* based models, best fit appears to be model 1b.

```{r compare1mods}
# compare models
compare(m1.1a, m1.1b, m1.1c)

```

Looks like the model with `temp_7_max`,`temp_30_mn`, and `temp_CV` performed best, but not significantly so. The model with `ppt` and other temp values may be important. Ultimately we need to make some plots to figure this out. Also helps to create an ensemble model so that any one parameter gets "averaged" and avoids issues with trying lots of different models until you get one that "works".

I can't figure out how to extract predictions from my ensemble model for a given set of variables, or how to make a residual plot. 

```{r ensemble_temp extraction, echo=F, eval=F}

# dummy data

d.pred <- data.frame(
    temp_7_max = seq(6,16, length.out=58),
    temp_CV = seq(1,100, length.out=58),
    temp_7_min = seq(6,16, length.out=58),
    breed = rbinom(58, 1, 0.5),
    temp_30_min = seq(6,16, length.out=58),
    temp_30_max = seq(6,16, length.out=58),
    temp_7_avg = seq(6,16, length.out=58)
)


# how the hell do I plot this madness?
m1_ens <- ensemble( m1.1a, m1.1b, data = d.pred)
mu <- apply( m1_ens$link , 2 , mean )
mu.PI <- apply( m1_ens$link , 2 , PI )

pch <- ifelse(d1$site=="NFA", 16, 1)
plot( d1$breed ~ d1$temp_7_max , col=rangi2 , pch=pch)

#plot(breed ~ temp_7_max, data=d1)
for ( i in 1:58 )
    abline( a=post1b$a[i] , b=post1b$bt7mx[i] ,
            col=col.alpha("black",0.3) )

lines(mu , d.pred$temp_7_max )
shade( mu.PI , d1$temp_7_max)


```

### Binomial: Varying Intercept Temp and Flow

Now add other variables into the model, with partial pooling (but no interactions):

```{r flowtemp varying intercept}

names(d1s)
summary(d1s)

# now just 7day flow and temp
m2.1a <- map2stan(
    alist(
        breed ~ dbinom( 1 , theta ) ,
        logit(theta) <- a + a_wy[WY] + bt7mx*temp_7_max + bt30mn*temp_30_min + 
          bdQ*deltQ + bQCV*Q_CV + bQmax*Q_max + bQ7CV*Q_7_CV + 
          bdLev*deltLev + blevCV*lev_CV,
        a ~ dnorm(1, 5),
        a_wy[WY] ~ dnorm( 0, sigma_WY ),
        c(bt7mx, bt30mn, bdQ, bQCV, bQmax, bQ7CV, bdLev, blevCV) ~ dnorm(0,1),
        sigma_WY ~ dcauchy(0,1)
    ), 
    data=d1s, warmup=4000, iter=1e4, chains=2, cores=2)

# view output
plot(m2.1a)
dev.off()
pairs(m2.1a)
dev.off()
precis(m2.1a, depth=2, warn = F)
plot(precis(m2.1a, depth=2))

# show density of log-odds
post2a <- extract.samples(m2.1a)
dens( post2a$bt7mx , ylim=c(0,1), xlim=c(-4,6))
dens( post2a$bdQ, col=rangi2 , lwd=2 , add=TRUE )
dens( post2a$blevCV, col="skyblue2" , lwd=2 , add=TRUE )
dens( post2a$bt30mn , col="red", add=TRUE)
dens( post2a$bdLev , col="green", add=TRUE)


coef(m2.1a)

# make some plots
plot(y=d1$breed, x=d1$temp_7_max, pch=ifelse(d1$site=="NFY", 16, 1), col=rangi2)
plot(y=d1$lev_CV, x=d1$temp_7_max, pch=ifelse(d1$site=="NFY", 16, 1), col=rangi2, ylim=c(0,20))
abline(a=coef(m2.1a)["blevCV"] , b=coef(m2.1a)["bt7mx"] , col=col.alpha("black",0.3))

for (i in 1:30)
  abline( a=post2a$a_wy[i], b=logistic(post2a$bt7mx[i]))


# compute expected value at MAP
mu <- coef(m2.1a)['a'] + coef(m2.1a)['a_wy[2]'] + coef(m2.1a)['bt7mx']*d1s$temp_7_max
# compute residual
m.resid <- d1s$temp_7_max - mu


plot( temp_7_max ~ lev_CV , d1s , pch=ifelse(d1s$river==1, 16, 1), col=rangi2, axes=F)
atx <- c(seq(-2, 3.1, 0.5))
aty <- c(seq(-2, 2.2, 0.5))
labels <- atx*sd(d1$lev_CV) + mean(d1$lev_CV)
labelsy <- aty*sd(d1$temp_7_max) + mean(d1$temp_7_max)
axis(side=1, at = atx, labels=round(labels, 1))
axis(side=2, at = aty, labels=round(labelsy, 1))
#abline( a=coef(m2.1a)['blevCV'], b=coef(m2.1a)['bt7mx']*d1s$temp_7_max[1])
abline( a=coef(m2.1a)['a'], b=coef(m2.1a)['bt7mx'])

for (i in 1:30)
  abline( a=post2a$a_wy[i], b=post2a$bt7mx[i], col=col.alpha("gray", alpha=0.6))


# loop over States
for ( i in 1:length(m.resid) ) {
    x <- d1s$lev_CV[i] # x location of line segment
    y <- d1s$temp_7_max[i] # observed endpoint of line segment
    # draw the line segment
    lines( c(x,x) , c(mu[i],y) , lwd=0.5 , col=col.alpha("black",0.7) )
}

```


#### Notes from Myfanwy 2017-05-24

Use DOWY, not binomial, sites are coin flips. What triggers breeding, but doesn't help answer more important factors of what drives breeding? Want response to be as continuous as possible. 

**DOWY as response**
gammaPoisson, has individual "rate", rate should be affected in same way by different predictor variables. Different sites as nested variables, rate varies by site. 

**Survival Analysis**

How long does an event take to occur, (i.e., time to breeding)? Freezing over in a lake as example. Continuous segmented regression, (tobit regression) + probit model with time series with latent variable.

**MultiState Model**

predicting frequencies that will be occupied on radio, occupied vs. not.

Hidden Markov model, model conditions leading up to observable state (hidden state). Hidden markov poisson, describe variability that characterizes this latent state that you don't observe, but you know it drives the latent state.

Eric Van Cleaves (multistate modeler, spatial component)

Input matrix: 365 x 30 variables
Each row is day of year
Whether a breeding event occurred on that day as binomial


## Continuous gamma-Poisson Model or Beta Binomial

pg350 in Rethinking: "*Predictor variables adjust the shape of this distribution, not the expected value of each observation.  The gamma-Poisson model is very much like a beta-binomial model, with the gamma distribution of rates (or expected values) replacing the beta distribution of probabilities of success.*"

```{r viewCurves}
# gammaPois
mu <- 3
theta <- 1
curve( dgamma2(x,mu,theta) , from=0 , to=10 )

#betaBinom
pbar <- 0.5
theta <- 5
curve( dbeta2(x,pbar,theta) , from=0 , to=1 ,
       xlab="probability" , ylab="Density" )

```

As theta approaches zero, the gamma distribution approaches a Gaussian distribution with the same mean value.  It will try to account for unobserved heterogeneity in probabilities and rates. Don't use WAIC!

**Don't use WAIC with gammaPoisson**
You should not use WAIC with these models, however, unless you are very sure of what you are doing.  e reason is that while ordinary binomial and Poisson models can be aggre- gated and disaggregated across rows in the data, without changing any causal assumptions, the same is not true of beta-binomial and gamma-Poisson models.  e reason is that a beta- binomial or gamma-Poisson likelihood applies an unobserved parameter to each row in the data. When we then go to calculate log-likelihoods, how the data are structured will deter- mine how the beta-distributed or gamma-distributed variation enters the model.

**Use DIC**

### Beta-Binomial
```{r}
library(rethinking)
names(d1)
d1$lagEM <- coerce_index(d1$lagEM)
d1$site <- coerce_index(d1$site)

m11.5 <- map2stan(
    alist(
        DOWY ~ dbetabinom(temp_7_max,pbar,theta),
        logit(pbar) <- a + b*lagEM + blevCV*lev_CV + bwyi*wyind,
        c(a, b, blevCV, bwyi) ~ dnorm(0,2),
        theta ~ dexp(1)
    ),
    data=d1,
    constraints=list(theta="lower=0"),
    start=list(theta=3),
    iter=4000 , warmup=1000 , chains=2 , cores=2 )
```

### stanarm?
```{r diff rstanpkg}

library(rstanarm)

d1b <- filter(d1, breed==1)
summary(d1b)

# try fitting a glm with interaction between best flow and the best temp metric
fit <- stan_glm(DOWY ~ temp_7_max * deltLev, data = d1b, family = Gamma, 
                prior_intercept = normal(0,1), prior = normal(0,1),
                chains = 4, cores = 2)

print(fit, digits = 3)
plot(fit)
pp_check(fit, plotfun = "stat_2d", stat = c("mean", "sd"))

dowy_SEQ <- seq(from = 200, to = 280, length.out = 20)
temp_SEQ <- seq(from = 6, to=16, length.out = 20)
y_nohs <- posterior_predict(fit, newdata = data.frame(temp_7_max = temp_SEQ, DOWY = dowy_SEQ, deltLev = -.01))
y_hs <- posterior_predict(fit, newdata = data.frame(temp_7_max = temp_SEQ, DOWY = dowy_SEQ, deltLev = -.1))
dim(y_hs)

par(mfrow = c(1:2), mar = c(5,4,2,1))
boxplot(y_hs, axes = FALSE, outline = FALSE, ylim = c(150,350),
        xlab = "Temp7mx", ylab = "Predicted DOWY", main = "deltLev = -0.1")
axis(1, at = 1:ncol(y_hs), labels = round(temp_SEQ, 0), las = 3)
axis(2, las = 1)
boxplot(y_nohs, outline = FALSE, col = "red", axes = FALSE, ylim = c(150,350),
        xlab = "Temp7mx", ylab = NULL, main = "delLev= -0.01")
axis(1, at = 1:ncol(y_hs), labels = round(temp_SEQ,0), las = 3)


```


```{r gamPoisson}

d1f <- d1 %>% filter(breed==1) # NFY and NFA breeding only

#change delts to positive:
d1f$deltQexp <- exp(d1f$deltQ)
d1f$deltLevexp <- exp(d1f$deltLev)
d1f$log_DOWY <- log(d1f$DOWY)

m3 <- map2stan(
  alist(
    log_DOWY ~ dgampois(mu, scale),
    log(mu) <- a + bt7mx*temp_7_max + bwyi*Index + bdeltLev * deltLevexp +
      bWairmx*W_air_30_max + blevCV*lev_CV,
    a ~ dnorm(0, 10),
    c(bt7mx, bwyi, bdeltLev, bWairmx, blevCV) ~ dnorm(0,1),
    scale ~ dcauchy(0,1)
  ),
  data=d1f,
  constraints=list(scale="lower=0"),
  start=list(scale=1),
  warmup=1000, iter=8000, chains=2, cores=2)

precis(m3)
plot(m3)

postcheck(m3)
#blue points show the empirical proportion on each row of the data
#open circles are the posterior mean p, with 89% percentile interval
# + symbols mark the 89% interval of predicted counts for breeding. Lots of dispersion expected here.


post3 <- extract.samples(m3)
quantile( logistic(post3$a) , c(0.025,0.5,0.975) )

# draw posterior mean dgam distribution
curve( dgampois(x,mean(logistic(post3$a)),mean(post3$scale)) , from=0 , to=1 ,
    ylab="Density" , xlab="probability", ylim=c(0,3) , lwd=2 )
# draw 100 beta distributions sampled from posterior
for ( i in 1:100 ) {
    p <- logistic( post3$a[i] )
    scale <- post3$scale[i]
    curve( dgampois(x,p,scale) , add=TRUE , col=col.alpha("black",0.2) )
}


```

### Multilevel Poisson

```{r}

m12.6 <- map2stan(
  alist(
    total_tools ~ dpois(mu),
    log(mu) <- a + a_society[society] + bp*logpop,
    a ~ dnorm(0,10),
    bp ~ dnorm(0,1),
    a_society[society] ~ dnorm(0,sigma_society),
    sigma_society ~ dcauchy(0,1)
),
data=d ,
iter=4000 , chains=3 )


post <- extract.samples(m12.6)
d.pred <- list(
    logpop = seq(from=6,to=14,length.out=30),
    society = rep(1,30)
)
a_society_sims <- rnorm(20000,0,post$sigma_society)
a_society_sims <- matrix(a_society_sims,2000,10)
link.m12.6 <- link( m12.6 , n=2000 , data=d.pred ,
    replace=list(a_society=a_society_sims) )

# plot raw data
plot( d$logpop , d$total_tools , col=rangi2 , pch=16 ,
    xlab="log population" , ylab="total tools" )
# plot posterior median
mu.median <- apply( link.m12.6 , 2 , median )
lines( d.pred$logpop , mu.median )
# plot 97%, 89%, and 67% intervals (all prime numbers)
mu.PI <- apply( link.m12.6 , 2 , PI , prob=0.97 )
shade( mu.PI , d.pred$logpop )



```



#### Flow and Stage

SHould think about varying intercept as water year or index and not just by river. 
```{r flow_temp interaction}

# filter to only breeding points and look at interactions between these two vars

d1f <- d1s %>% filter(breed==1) # NFY and NFA breeding only
names(d1f)
d1f$deltQexp <- exp(d1f$deltQ)
d1f$deltLevexp <- exp(d1f$deltLev)
d1f$exp_DOWY <- exp(d1f$DOWY)

# now just 7day temp + lev_CV

m2.1b <- map2stan(
    alist(
        log_DOWY ~ dnorm( mu , sigma ) ,
        mu <- a + a_wy[WY] + bt7mx*temp_7_max + bdeltLev*deltLevexp +
          bWairmx*W_air_30_max,
        c(a, bt7mx, bdeltLev, bWairmx) ~ dnorm(0,1),
        a_wy[WY] ~ dnorm( 0, sigma_WY),
        sigma ~ dcauchy(0,1),
        sigma_WY ~ dcauchy(0,1)
    ), 
    data=d1f, warmup=1000, iter=5000, chains=4, cores=2)

# view output
plot(m2.1b)
dev.off()
pairs(m2.1b)
dev.off()
precis(m2.1b, depth=2, warn = F)
plot(precis(m2.1b, depth=2))

# show density of log-odds
post2b <- extract.samples(m2.1b)
dens( post2b$bdeltLev, col="skyblue2" , lwd=2 )
coef(m2.1b)

# make some plots
plot(y=d1sB$lev_CV, x=d1sB$temp_7_max, pch=ifelse(d1$site=="NFY", 16, 1), col=rangi2)
abline(a=coef(m2.1b)["a"] , b=coef(m2.1b)["blevCV"] , col="black", lty=2)

for (i in 1:12)
  abline( a=post2b$a_wy[i], b=post2b$blevCV[i])


# compute expected value at MAP
mu <- coef(m2.1b)['a'] + coef(m2.1b)['blevCV']*d1sB$lev_CV
# compute residual
m.resid <- d1sB$lev_CV - mu


plot( temp_7_max ~ lev_CV   , d1sB , pch=ifelse(d1s$WY==1, 16, 1), col=rangi2, axes=F)
atx <- c(seq(-2, 3.1, 0.5))
aty <- c(seq(-2, 2.2, 0.5))
labelsx <- atx*sd(d1$lev_CV) + mean(d1$lev_CV)
labelsy <- aty*sd(d1$temp_7_max) + mean(d1$temp_7_max)
axis(side=1, at = atx, labels=round(labelsx, 1))
axis(side=2, at = aty, labels=round(labelsy, 1))
abline( a=coef(m2.1b)["a"] , b=coef(m2.1b)["blevCV"]*d1sB$lev_CV)

# loop over States
for ( i in 1:length(m.resid) ) {
    x <- d1sB$lev_CV[i] # x location of line segment
    y <- d1sB$temp_7_max[i] # observed endpoint of line segment
    # draw the line segment
    lines( c(x,x) , c(mu[i],y) , lwd=0.5 , col=col.alpha("black",0.7) )
}
```

```{r jacobgampoisson}

#########
# from Jacob 05/17/17
mz43 <- map2stan(
  alist(
    Zoops ~ dgampois(mu,scale),
    log(mu) <- a_site[site_id] + bt*temp + bno * NO3.N + bp * PO4.P + bd * DOC + bj * JulianDay,
    a_site[site_id] ~ dcauchy(10,sigma_site),
    sigma_site ~ dcauchy(1,1),
    c(bt,bno,bp,bd,bj) ~ dnorm(0,1),
    scale ~ dcauchy(0,2)
  ),
  data=zb,
  constraints=list(scale="lower=0"),
  start=list(scale=2)
)

```


### Geometric Distribution

The geometric distribution might make some sense (Ch 10, pg 328), as it describes the number of events up until something happens (such as an event, like the start of breeding). If we want to know the probability of that event (often called *Event History Analysis* or *Survival Analysis*), we can use the `dgeom` distribution in R for our common likelihood function. 

The only caveat is that we need to **assume** the probability of the terminating event is constant through time (or distance), and the units of time (or distance) are discrete.

```{r geometric distrib}
# simulate
N <- 100
x <- runif(N)
y <- rgeom( N , prob=logistic( -1 + 2*x ) )

# estimate
m10.18 <- map(
    alist(
        y ~ dgeom( p ),
        logit(p) <- a + b*x,
        a ~ dnorm(0,10),
        b ~ dnorm(0,1)
),
    data=list(y=y,x=x) )
precis(m10.18)
plot(precis(m10.18))

```

## LASSO glm 

```{r lasso}

library(glmnet)

# Only UNREG
d.1 <- dplyr::select(rabo_logist, DOWY, everything(), -date, -EM_per_km, -REG,
                     -(obs_strt:apr_jul), -len_km, -starts_with("CDEC"), 
                     -station, -DOY, -WYsum, -lev_7_avg, -Q_cfs, 
                     -Q_min, -Q_7_cfs) %>% 
  filter(site=="NFY" | site=="NFA") %>% 
  as.data.frame

# RM NAs, add id and site to rownames:
d.1 <- d.1 %>% filter(!is.na(temp_30_min))
d.1_rownames<- paste0(d.1$site, "-",d.1$WY, "-", seq(1:nrow(d.1)))
d.1 <- select(d.1, -site, -WY)
d.1$breed <-as.factor(d.1$breed)
d.1$lagEM <- as.factor(d.1$lagEM)

# make the model matrix for binomial
x <- model.matrix(breed~., data=d.1)
x <- x[,-1] # remove the response var column

# model matrix for gaussian
xG<- d.1 %>% filter(lagEM=="d0") %>% select(-breed, -lagEM)
x2 <- model.matrix(DOWY~., data=xG)
x2 <- x2[,-1] # remove the response var column

# MODEL: BINOMIAL
fit = glmnet(x=x, y=d.1$breed, family = "binomial")
plot(fit, label=T)
print(fit)
coef(fit,s=0.05)

# MODEL: GAUSSIAN
fit2 = glmnet(x=x2, y=xG$DOWY)
plot(fit2, label=T)
print(fit2)
coef(fit2,s=0.05)

# create matrix for prediction (rows, cols)
nx = matrix(rnorm(nrow(x2)*ncol(x2)),nrow(x2),ncol(x2))
predict(fit2,newx=nx,s=c(0.1,0.05)) # new predictions

# cross validated vs. breed
cvfit <- cv.glmnet(x=x2, y=xG$DOWY)
plot(cvfit)
cvfit$lambda.min
coef(cvfit, s=0.05)
coef(cvfit, s="lambda.min")

```

lamda.min is the value of 位位 that gives minimum mean cross-validated error. The other 位位 saved is lambda.1se, which gives the most regularized model such that error is within one standard error of the minimum. To use that, we only need to replace lambda.min with lambda.1se above

```{r}

newMod <-predict(cvfit, newx = x2, s = "lambda.min") %>% data.frame()
colnames(newMod)<-"predictions"
newMod$siteID<- d.1_rownames[1:12]

ggplot(newMod) + geom_point(aes(x=siteID, y=predictions), pch=16) +
  coord_flip() + ylab("Prob of Breeding") + xlab("")

```


## Logistic `glm` Model

Now let's run a model. See [here](http://stats.idre.ucla.edu/r/dae/logit-regression/)

```{r model}

# Only UNREG
xG<- d.1 %>% filter(lagEM=="d0") %>% select(-breed, -lagEM)

# All sites
d.2 <- dplyr::select(rabo_logist, DOWY, everything(), -date, -Index, -WYsum,
                     -EM_per_km, -station, -(obs_strt:apr_jul), 
                     -len_km, -starts_with("CDEC"), -DOY, -WYsum,
                     -Q_min, -Q_7_cfs, -lev_7_avg) %>%
  as.data.frame

# fix index
#d.2$breed <-as.integer(as.factor(d.2$breed))
d.2$lagEM <- as.integer(as.factor(d.2$lagEM))
d.2$site <- as.integer(as.factor(d.2$site))
d.2$REG <- as.integer(as.factor(d.2$REG))

# scaled
d.2.s<-scale(d.2) %>% as.data.frame


# run a gaussian response model (unscaled)
mod.gauss <- glm(DOWY ~ deltQ + Q_CV + Q_7_CV +
                   W_air_7_avg + W_air_30_max + W_humidity_avg +
                   temp_7_avg + temp_7_max + temp_30_min + temp_30_max +
                   temp_CV + lev_CV + deltLev + Index +
                   days_no_ppt, data = xG, family = "gaussian")
summary(mod.gauss)
anova(mod.gauss)

# run a logistic binomial response model (unscaled)
mod.logit <- glm(breed ~ deltQ + Q_CV + Q_7_CV +
                 W_air_7_avg + W_air_30_max + W_humidity_avg +
                 temp_7_avg + temp_7_max + temp_30_min + temp_30_max +
                 temp_CV + lev_CV + deltLev +
                 days_no_ppt + site, data = d.2, family = "binomial")
summary(mod.logit)
anova(mod.logit)
```

So **Deviance residuals** are a measure of model fit, and `AIC = 178.8`. For every one unit change in **`temp_7_max`** there is a log-odds of oviposition timing (breeding) increase of 20.2224, whereas a change in the stage (`lev_CV`) yields a strong negative change in the log odds of oviposition by -2.09

You can calculate the confidence intervals using the `confint` function.

```{r}
confint(mod.logit)

```

```{r}
# simulate and predict

newdata1 <- d %>% select(DOWY, site, lev_avg:deltQ) %>% group_by(site,  DOY) #%>% 
  summarize(airmin.7 = mean(airmin.7),
            days_wo_ppt = mean(days_wo_ppt),
            level.avg = mean(level.avg),
            temp.min.7 = mean(temp.min.7),
            level.7dL = mean(level.7dL))

# predict
newdata1$rankP <- predict(mylogit, newdata = newdata1, type = "response")
newdata1
ggplot() + geom_point(data=newdata1, aes(x=DOY, y=rankP, color=as.factor(river_id), shape=as.factor(WY_id)), size=3)
```




## BRT

```{r brt}
library(gbm)
library(dismo)
library(ggplot2)
library(viridis)

set.seed(33)

# gbm First Model -----
d.1brt <- dplyr::select(rabo_logist, DOWY, everything(), 
                     -date, -EM_per_km, -REG,
                     -(obs_strt:apr_jul), -len_km,
                     -starts_with("CDEC"), 
                     #-starts_with("W_"), 
                     -station, -DOY, -WYsum, 
                     -lev_7_avg, -Q_cfs, 
                     -Q_min, -Q_max, -Q_7_cfs) %>% 
  filter(site=="NFY" | site=="NFA") %>% 
  filter(!is.na(temp_30_min)) %>% 
  dplyr::select(breed, everything(), -Index, -lev_avg, -DOWY) %>%
  as.data.frame()
d.1brt$site <- as.factor(d.1brt$site)
d.1brt$lagEM <- as.factor(d.1brt$site)

gbm3s <- gbm.step(data = d.1brt,
                     gbm.x = 2:26,    
                     gbm.y = 1,
                     family = "bernoulli",  
                     tree.complexity = 2,   
                     learning.rate = 0.001, 
                     bag.fraction = 0.75,   
                     n.folds = 5,      
                     plot.main=T,
                     verbose=T) 

paste0("mean estimated deviance from CV: ", round(gbm3s$cv.statistics$deviance.mean,3))
paste0("SE estimated deviance from CV: ", round(gbm3s$cv.statistics$deviance.se,3))

# gbm simplify barplot ----

par(mar=c(5,12,3,3))
barplot(rev(summary(gbm3s, plotit=FALSE)$rel.inf), 
        horiz = TRUE, col = viridis(length(gbm3s$var.names)), 
        names = rev(summary(gbm3s, plotit=FALSE)$var), 
        xlab = "Relative influence",
        las=1, 
        main="Relative Influence")

topVars<-summary(gbm3s, plot=F) %>% as.tibble %>% filter(rel.inf>4)
topVars


## Another BRT but simplified
names(d.1brt)
topVars$var <- factor(topVars$var)
d.1simp <- dplyr::select(d.1brt, breed, site, deltQ, W_air_30_max,
                         temp_30_min, temp_7_max, deltLev,
                         W_humidity_avg, lev_CV)

gbm3s <- gbm.step(data = d.1simp,
                     gbm.x = 2:9,    
                     gbm.y = 1,
                     family = "bernoulli",  
                     tree.complexity = 2,   
                     learning.rate = 0.001, 
                     bag.fraction = 0.75,   
                     n.folds = 5,      
                     plot.main=T,
                     verbose=T) 

paste0("mean estimated deviance from CV: ", round(gbm3s$cv.statistics$deviance.mean,3))
paste0("se estimated deviance from CV: ", round(gbm3s$cv.statistics$deviance.se,3))

# gbm simplify barplot ----

par(mar=c(5,12,3,3))
barplot(rev(summary(gbm3s, plotit=FALSE)$rel.inf), 
        horiz = TRUE, col = viridis(length(gbm3s$var.names)), 
        names = rev(summary(gbm3s, plotit=FALSE)$var), 
        xlab = "Relative influence",
        las=1, 
        main="Relative Influence")

summary(gbm3s, plot=F)
```

