---
title: "Oviposition Plasticity"
author: "Ryan Peek"
date: "Updated: `r format(Sys.Date())`"
output: html_document
---

```{r setup, include=FALSE, echo=F}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = normalizePath("../"))

# make things beep on error and finish:
options(error = function(){    # Beep on error
  beepr::beep()
  Sys.sleep(1)
  }
 )

.Last <- function() {          # Beep on exiting session
  beepr::beep(2) # beep(8) is awesome but too long
  Sys.sleep(1)
}

# if you want to reset these so they don't run: options(error=NULL)

```

## What Permits Plasticity?

How might variable environmental conditions provide cues for breeding timing from year to year? For frogs that have an amazing plasticity in breeding oviposition timing (inital deposition of eggs) such as *Rana boylii*, there must be certain thresholds that exist (a cutoff which makes oviposition not possible, physiologically). However, beyond a given threshold, there also may be certain cues which act as important environmental markers in dynamic environments such as rivers. By utilizing both, it is likely that the odds of reproductive success improve, and over time this results in evolutionary success. For example, frogs that breed to soon are more likely to have eggs scoured out by late spring storms or have longer developmental/hatching times because water temperatures are lower. Frogs breeding later in the season may trade off greater egg deposition and hatching success with later metamorphosis and lower overwinter survival. 

We know that frogs from the same populations in the exact same locations can deposit eggs more than a month later or many weeks earlier than the average observed date of deposition, depending on the annual conditions (i.e., wetter or drier years). We suspect that the water temperature may be the most important driver in this plasticity, as it may act as a threshold which once exceeded, provides a very strong indicator of "summer" or future hydrologic condition.

However, the flow in a given reach of river can also be highly correlated with water temperature. Linkage between water temperature and patterns in flow may be tightly tied, particularly in rivers in the Sierra Nevada. Historically, Sierra Nevada river hydrology has been driven by a general Mediterranean pattern of wet (though variable) winter weather followed by spring snowmelt and warmer and drier summers. While temperature may exist as a spawning threshold, the actual initiation of breeding or spawning is highly variable in **R. boylii**, and likely requires additional enviromental cues such as changing water levels, flow magnitude, air temperature, humidity, etc.

These additional metrics may each act as environmental cues, but to conserve plasticity across environmental variablity present in a wide geographic range, it is unlikely that any of these can dictate oviposition independently (whereas a water temperature threshold may be a sole driver in some years). In order to determine the strongest predictor of breeding, we've collected oviposition data over 6 rivers in the Sierra Nevada from 2011 through 2016. In addition we collected a variety of metrics related to flow, air temperature, water temperature, water year index, day of water year, precipitation, and barometric pressure.

Using two different methods, we compare the variation explained by these metrics. PCA and BRT. 


### The Data
```{r get data}
library(tidyverse)
library(lubridate)

#sites <- c("NFA", "NFY") # if you want to restrict the sites

frogBreed <- read_csv("data/oviposition_start_mainstem_sites.csv") %>% 
  mutate(estim_strt=mdy(estim_strt),
         obs_strt=mdy(obs_strt)) %>%
  select(Site:WYT, REG:totalEM)
  #filter(Site %in% sites)

hydroDat <- read_csv("data/oviposition_glm_stgtemps_long.csv") %>% 
  mutate(Date=mdy(Date),
         site=as.factor(toupper(site)))

# add WY Index
wys<-read_csv(file = "data/cdec_wy_index.csv") %>% filter(Basin=="SAC")

```

### Tidy and Merge the Data

```{r merge_data}
source('scripts/functions/f_doy.R') # for adding year date cols

# join the data together:
data1 <- left_join(hydroDat, frogBreed, by = c("Date"="estim_strt", "site"="Site")) %>% 
  mutate(site = as.factor(site))
data1 <- add_WYD(data1, "Date") %>% select(-WYT) %>% 
  mutate(DOY = as.integer(DOY),
         WY = as.integer(WY),
         DOWY = as.integer(DOWY))

# join with WY index data
data1 <- inner_join(data1, wys[,c(1,4:5)], by="WY")

names(data1)

# select only data between May-Jun 
df <- data1 %>% #filter(site %in% sites) %>% 
  filter(DOY>120, DOY<182) %>% 
  select(Date, site, obs_strt, totalEM, DOY:DOWY, air.7:level.7dL, WYsum, Index, -ppt2_in, -days_wo_ppt2)

summary(df)

# remove NAs:
df2 <- df %>% 
  filter(!is.na(air.7), 
         !is.na(level.avg))
summary(df2)

# add binary 1 or 0 for breeding
df2 <- df2 %>% mutate(ovipos=as.integer(ifelse(is.na(totalEM), 0, 1)))
names(df2)

# get data
d1 <- df2 %>% dplyr::select(ovipos, site, WYsum, Index, DOWY, air.7:level.7dL) %>% 
  filter(!is.na(ppt_in))
summary(d1)

# refactor the sites to drop unused factors
d1$site <- factor(d1$site)

# drop factors
d2 <- dplyr::select(d1, -site) %>% setNames(., gsub("\\.", "_", colnames(.))) %>% as.data.frame
names(d2)
```

## PCA

Try a PCA of the variables to see which is strongest predictor and potentially reduce collinearity.

```{r pca}
# filter to only the breeding points
bdat<-filter(d2, ovipos==1) %>% row.names() %>% as.integer()
traindat <- d2[bdat,]

#remove the dependent and identifier variables
traindat <- traindat %>% select(-ovipos, -ppt_in, -WYsum)

# PCA
prin_comp <- prcomp(traindat)
prin_comp$center

# rotation: The rotation measure provides the principal component loading. Each column of rotation matrix contains the principal component loading vector. This is the most important measure we should be interested in
prin_comp$rotation

# summary of components
summary(prin_comp) # so PC1 : PC3 explain 83% of the total variance in the data

# loadings per variable PC1
prin_comp$rotation[,1:3] %>% as.data.frame %>%
  mutate(vars = row.names(prin_comp$rotation)) %>% 
  select(vars, PC1) %>% arrange(PC1)

# loadings per variable PC2 & PC3
prin_comp$rotation[,2:3] %>% as.data.frame %>%
  mutate(vars = row.names(prin_comp$rotation)) %>% 
  select(vars, PC2) %>% arrange(PC2)

# loadings per variable PC3
prin_comp$rotation[,1:3] %>% as.data.frame %>%
  mutate(vars = row.names(prin_comp$rotation)) %>% 
  select(vars, PC3) %>% arrange(PC3)


# plot
biplot(prin_comp, choices = 1:2, scale = 0, pc.biplot = F, xlabs=rep(".", nrow(traindat)))
biplot(prin_comp, choices = 2:3, scale = 0, pc.biplot = F, xlabs=rep(".", nrow(traindat)))
dev.off()

# a quad plot of results

# THE FUNCTION:
pcaCharts <- function(x) {
    x.var <- x$sdev ^ 2
    x.pvar <- x.var/sum(x.var)
    print("proportions of variance:")
    print(x.pvar)
    
    par(mfrow=c(2,2))
    plot(x.pvar,xlab="Principal component", ylab="Proportion of variance explained", ylim=c(0,1), type='b')
    plot(cumsum(x.pvar),xlab="Principal component", ylab="Cumulative Proportion of variance explained", ylim=c(0,1), type='b')
    screeplot(x)
    screeplot(x,type="l")
    par(mfrow=c(1,1))
}

# plot
pcaCharts(prin_comp)

```

```{r ggbiplots}
# library(devtools)
# install_github("ggbiplot","vqv")
library(ggbiplot)

g <- ggbiplot(prin_comp, obs.scale = 1, var.scale = 1,
              ellipse = TRUE, 
              circle = TRUE)
g <- g + scale_color_discrete(name = '')
g <- g + theme(legend.direction = 'horizontal', 
               legend.position = 'top')
print(g)

# Another PCA option
library(caret)

# remove the categorical vars from data
bdat <- filter(d2, ovipos==1) %>% select(-ppt_in)
summary(bdat)

df.cont <- as.data.frame(bdat[,-1])
df.cat <- as.data.frame(bdat[,1])

# preprocess with caret
df.trans <- preProcess(x = df.cont, method = c("BoxCox","center","scale"))
df.trans

# predict on data
df.preproc <- predict(df.trans,newdata = df.cont)

# PCA
df.pca <- prcomp(df.preproc, center = FALSE)
df.pca

# loadings per variable PC1
df.pca$rotation[,1:3] %>% as.data.frame %>%
  mutate(vars = row.names(df.pca$rotation)) %>% 
  select(vars, PC1) %>% arrange(PC1)

# loadings per variable PC2 & PC3
df.pca$rotation[,2:3] %>% as.data.frame %>%
  mutate(vars = row.names(df.pca$rotation)) %>% 
  select(vars, PC2) %>% arrange(PC2)

# loadings per variable PC3
df.pca$rotation[,1:3] %>% as.data.frame %>%
  mutate(vars = row.names(df.pca$rotation)) %>% 
  select(vars, PC3) %>% arrange(PC3)

pcaCharts(df.pca)

# now switch off scaling factors (var.scale)
g <- ggbiplot(df.pca, choices = 2:3, scale = 0, var.scale = 0, labels= df.cat$site, groups = df.cat$site,
              ellipse = TRUE, 
              circle = TRUE)
g <- g + scale_color_discrete(name = '')
g <- g + theme(legend.direction = 'horizontal', 
               legend.position = 'top')
print(g)

```





## BRT Model

```{r brt1}
library(dismo)

# get data
d <- nfdf %>% dplyr::select(ovipos, site,WY, DOWY, air.7:level.7dL) %>% 
  filter(!is.na(ppt_in))
summary(d)

# fix indexes
d$river_id <- as.integer(as.factor(d$site))
d$WY_id <- as.integer(as.factor(d$WY))
d$DOWY <- as.integer(d$DOWY)

# drop factors
d <- dplyr::select(d, -site, -WY) %>% setNames(., gsub("\\.", "_", colnames(.))) %>% as.data.frame
names(d)

egg.tc5.lr01 <- gbm.step(data= d, gbm.x = 2:23, gbm.y = 1,
                         family = "bernoulli", tree.complexity = 4,
                         learning.rate = 0.0005, bag.fraction = 0.5)

summary(egg.tc5.lr01)

# plot
gbm.plot(egg.tc5.lr01, rug = T, n.plots = 12, write.title = FALSE)
gbm.plot.fits(egg.tc5.lr01)

# interactions
find.int <- gbm.interactions(egg.tc5.lr01)
find.int$interactions
find.int$rank.list

# plot pairwise interactions:
gbm.perspec(egg.tc5.lr01, 7, 5)#y.range=c(15,20), z.range=c(0,0.6))
```


```{r brt2}

library(dismo)
tst<- read.csv("./data/HSC/egg_use_availability_equal_samples.csv")
tst2<- read.csv("./data/HSC/Eggs_HSC_ncss_export.csv") %>% 
  mutate(Date= as.Date(Date,  origin="1899-12-30", tz="America/Los_Angeles")) %>% 
  select(-starts_with("Random"))
  #mutate(Date= as.POSIXct(Date * (60*60*24),  origin="1899-12-30", tz="America/Los_Angeles")) # for datetime
head(tst2)

head(tst)
tst1 <- tst %>% sample_n(size=200, replace=FALSE) # create training set

egg.tc5.lr01 <- gbm.step(data=tst1, gbm.x = 3:5, gbm.y = 1,
                         family = "bernoulli", tree.complexity = 5,
                         learning.rate = 0.01, bag.fraction = 0.5)

summary(egg.tc5.lr01)

# plot
gbm.plot(egg.tc5.lr01, rug = T, n.plots = 8, write.title = FALSE)
gbm.plot.fits(egg.tc5.lr01)

# interactions
find.int <- gbm.interactions(egg.tc5.lr01)
find.int$interactions
find.int$rank.list

# plot pairwise interactions:
plot_ly(gbm.perspec(egg.tc5.lr01, 1, 2))#y.range=c(15,20), z.range=c(0,0.6))

# alternate version with use vs. available
tst2$UseBinary <- ifelse(tst2$Data_Type=="USE", 1, 0)
tst2_brt <- select(tst2, UseBinary, 10, 5:8, 11:23) %>% 
  filter(!is.na(Sub_Cont_6),
         !Meso_Type=="")

egg2.tc5.lr01 <- gbm.step(data=tst2_brt, gbm.x = 2:19, gbm.y = 1,
                         family = "bernoulli", tree.complexity = 5,
                         learning.rate = 0.01, bag.fraction = 0.5)

summary(egg2.tc5.lr01)

# plot
gbm.plot(egg2.tc5.lr01, rug = T, n.plots = 12, write.title = FALSE)
gbm.plot.fits(egg2.tc5.lr01)

# interactions
find.int <- gbm.interactions(egg2.tc5.lr01)
find.int$interactions
find.int$rank.list

# plot pairwise interactions:
gbm.perspec(egg2.tc5.lr01, 5,4)#y.range=c(15,20), z.range=c(0,0.6))


```

## Logistic `glm` Model

Now let's run a model. See [here](http://stats.idre.ucla.edu/r/dae/logit-regression/)

```{r model}

# get data
d <- nfdf %>% select(ovipos, site, WY, DOWY, air.7:level.7dL)

# fix index
d$river_id <- as.integer(as.factor(d$site))
d$WY_id <- as.integer(as.factor(d$WY))

xtabs(~ovipos + DOY, data = d)
xtabs(~ovipos + days_wo_ppt, data = d)

# factor the data that is factorable
str(d)
d$site <- factor(d$site)
d$WY <- factor(d$WY)

# run a logistic binomial response model
mylogit <- glm(ovipos ~ river_id + WY_id + DOY + airmin.7 + days_wo_ppt + 
                 level.avg + temp.min.7 + level.7dL, data = d, family = "binomial")
summary(mylogit)

```

So **Deviance residuals** are a measure of model fit, and `AIC = 80.093`. For every one unit change in **`airmin.7`** there is a log-odds of oviposition (breeding) increase of 0.91, whereas a change in the water temperature (`temp.min.7`) yields a strong negative change in the log odds of oviposition by -1.114.

You can calculate the confidence intervals using the `confint` function.

```{r}
confint(mylogit)

library(aod)
# Can use a wald.test to test coefficients in order they appear in model
aod::wald.test(b= coef(mylogit), Sigma = vcov(mylogit), Terms=c(1,4))

## odds ratios and 95% CI
exp(cbind(OR = coef(mylogit), confint(mylogit)))

```


